Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-06 chapters/paper3.tex
Changes 1. Updated text for rheology of bijels section

While the microstructure and synthesis techniques for bijels have been explored in earnest, the rheology of bijels
has remained relatively unexplored. Some interesting phenomena for bijels and shearing include the 
presence of monogels, made by remixing the phase separated liquid domains of the bijel while the electrostatic
interactions between particles maintains the structure of the particle monolayer. \cite{sanz_colloidal_2009} 
Monogel formation has been observed in lutidine/water and styrene/butadiene oligomer based bijels but not ethane-diol nitromethane based bijels
\cite{sanz_colloidal_2009, bai_dynamics_2015, tavacoli_novel_2011} Computational Investigations have identified that confined bijels under shear undergo elongation of
the domains in the direction of shear succeeded by particle detachment from the interface and eventual failure of the bijel. \cite{bonaccorso_shear_2020}
More recently, bijels have been identified to be 2D glasses percolating in 3D space, characterized through comparing the complex rheology of bijels against
colloidal gels made from silica particles with electrostatic interactions. \cite{ching_bijel_2022}

In many of the manufacturing techniques outlined the rheological properties are essential in ensuring that 
the casting mixture remains processable. \cite{haase_continuous_2015,haase_situ_2016} In STrIPS, the flow rate of the casting mixture and the co-solvent
control the microstructure of the bijel and the morphology of the resulting bijel material, controlling whether droplets or ropes are synthesized. \cite{haase_continuous_2015}
When investigating the effect of magnetic fields on the rheological properties of suspensions, it has been shown that the viscosity of ferrofluids increases substantially due
to the self assembly of particles restricting the flow of fluid, greatly increasing the viscosity of the ferrofluid. cite{qiao_magnetorheological_2012} The application
of shear onto suspensions can also cause shear banding as interparticle interactions mediate particle fluid interactions, creating localized velocity differences to the bulk.
\cite{xu_relation_2013} Looking into emulsions, the application of magnetic fields increases the viscosity of emulsions, making them more shear thinning than when no field 
is applied.

In past chapters, we identified the microstructure and particle arrangement of bijels stabilized by ellipsoidal particles under magnetic fields and demonstrated
particles morphology specific ordering to the applied magnetic field. We also show that the local ordering of the particles is affected from the application of
the magnetic field. In this chapter, We probe the dynamics of bijels under constant shear to understand how these microstructural and particle monolayer changes
affect the rheology of bijels stabilized by ellipsoidal particles under magnetic fields. First, we define a shear capillary number, $Ca_s$
which in the literature lies 0.04 and 0.16. \cite{frijters_effects_2012, yang_capillary_2022}
However, these capillary number ranges would exceed the 
maximum mach number the model allows if this same range were used $(Ma \leq 0.03)$. To accommodate this limitation the 
largest capillary number used will be $Ca_s = 10^-4$ which correspond to a maximum of $Ma \approx 0.02$. As all particles used 
have only hard-sphere type interactions, it is expected that the behaviour seen should mimic 2D colloidal glasses 
percolating in 3D space, akin to what Ching and Mohraz saw, with an additional dependence on the direction of shear. 
This would predict the discovery of particle monolayer dependent elasticity and yield stress along with shear thinning 
behavior of the bijels.

To verify these predictions, bijel microstructure will be defined using four processing histories; The first is of a 
bijel simulated under a $\Bar{B} = 1$ magnetic field strength, the second is a bijel stimulated under no field, 
followed by the application of a $\Bar{B} = 1$ field after jamming, the third is a bijel simulated under no field, 
while the final microstructure is a bijel simulated under $\Bar{B} = 1$ magnetic field, followed by switching the 
field off after jamming. This gives insight into the impact of processing history on the shear properties of a bijel, 
in addition to the microstructural and colloidal insights gained. Based on the results in Bonaccorso et al., there 
should be shear driven and shear rate dependent elongation of the domains in the direction parallel to the applied 
shear which in this system will be seen as a reduction in $L_{\perp}$ and an increase in $L_{\parallel}$. 
\cite{bonaccorso_shear_2020} The microstructure anisotropy will also be a factor in the viscosity results, as 
larger domains are more permeable than smaller ones, meaning that bijels where $L_{\perp} > L_{\parallel}$ should 
see less of the domain elongation effects shown in Bonaccorso et al. as the permeability of the bijels rises with 
larger domain size. \cite{bonaccorso_shear_2020}

It is also expected that the effective viscosity will be different between the four microstructures dependent upon the 
degree of nematic ordering and microstructural anisotropy of the bijel. Nematic ordering of the particles in the direction of shear
should resulting in a lowered effective viscosity compared to bijels without nematically 
ordered particles. \cite{xu_relation_2013, vermant_flow-induced_2005} Tracking of the proportion of particles on the 
interface will also yield insight into how the packing of the particles affects the rate at which particles will get 
ejected from the interface. Bonnacorso et al. identified that shear applied onto bijels caused particles to align to the direction of
shear before being ejected from the interface. Systems with a larger $\eta_{eff}$ are predicted to have the largest domains, largest 
difference between initial and final particle order and lowest number of particles left on the interface once steady 
state has been established.

@@

We begin this investigation by analyzing the rheological response of bijels stabilized by ellipsoidal particles under constant shear.
Specifically, we investigate the effect that particle order has on the rheological response characterized. We are also interested in
identifying the impact that the application of magnetic fields have on the rheological response of the bijel. 
Bijel templates simulated in Aim 1 with no field and a field strength of $\bar{B} = 1$ stabilized with a prolate and oblate particles with a 
particle volume fraction $\phi_p = 0.15$ were used as the starting point for these simulations. Next, we apply a field strength of
$\bar{B} = 1$ or switch off the applied field to all structures. This results in four processing setups; $\bar{B}:0\rightarrow 0$, $\bar{B}:0\rightarrow 1$,
$\bar{B}:1\rightarrow 0$ and $\bar{B}:1\rightarrow 1$. We then define a shear capillary number,  $Ca_s = \frac{\dot{\gamma} R_{s} \eta_{f}}{\sigma}$ where 
$\dot{\gamma}$ is the applied strain rate, $R_s$ is the equivalent spherical radius of the particles, $R_s = 7.9$, $\eta_f$ is the dynamic viscosity of the 
fluid and $\sigma$ is the surface tension of the interface. We use five shear capillary numbers, 
$Ca_s = 1\cdot10^{-4}, 5\cdot10^{-5}, 1\cdot10^{-5}, 5\cdot10^{-6}, 1\cdot10^{-6}$.

@@

The ordering of the particles play an important role in determining the rheology of suspensions as ordered particles can slip over one another
before the structure breaks. The adsorption of particles on the interface of a bijel could affect how this slip occurs. Previous work investigating
the rheology of bijels have determined shear thinning behavior characterized with a Herschel-Buckley fit equation. \cite{macmillan_rheological_2019} 
After applying shear on the bijel templates, we characterize and plot the time evolution of the the shear stress in Figure \ref{fig:stress_time}.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/stress-time_compare.png} 
    \caption{Time evolution of the shear stress of bijels stabilized with ellipsoidal particles as a function of the applied strain rate as
             a function of the initial microstructure and applied magnetic field. The observed stress response is difficult to say when a steady
             state is reached.} 
    \label{fig:stress_time} 
\end{figure}

Figure \ref{fig:stress_time} shows that there is no definitive steady state value of the shear stress obtained in the simulations performed. This is
attributed to the coarsening of the domains and reorientation of particles at the interface. \cite{tavacoli_novel_2011,macmillan_rheological_2019} 
Tavacoli et al. investigated the rheological response of Ethanediol/Nitromethane bijels and identified that upon placement of a needle to calculate the yield
stress of the bijel, the structure of the material was irreversibly altered. \cite{tavacoli_novel_2011} Macmillian et al. identified that repeated 
shearing of a bijel synthesized using direct mixing resulted in degraded mechanical performance. \cite{macmillan_rheological_2019}
To characterize the fit, we identify pseudo-steady state regions within each of the shear stress plots and fit the stress identified within this 
pseudo-steady state region to the strain rate. Pseudo-steady state regions are identified by measuring where the rate of change of shear stress is 
smallest for the longest duration. We fit the calculated stress to the applied strain rate using a Herschel-Buckley rheology model defined as
$\sigma_{xz} = \sigma_{y} + K(\dot{\gamma})^{n}$. $\sigma_{xz}$ is the shear stress calculated from the simulation, $\sigma_{y}$ is the yield stress of the bijel, 
$K$ is the flow consistency index and $n$ is the flow index. Where appropriate, we include results corresponding to a shear capillary number of 
$Ca_s = 10^{-6}$ as in many stress responses, $Ca_s = 10^{-6}$ may not always yield usable results.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.4]{../figures/results/paper3/stress_strain-all.png} 
    \caption{Fitting the experimental shear stress to the strain rate using the Herschel Buckley model. We characterize the bijels are 
             shear thinning in all cases and that the bijels become less shear thinning as the particle order increases. We also
             demonstrate that the particle ordering affects the yield stress differently for particles stabilized by oblate and 
             prolate particles.} 
    \label{fig:stress_strain} 
\end{figure}

From Figure \ref{fig:stress_strain}, the slopes of the curve indicate a non-linear shear response to the applied strain rate, suggesting that the rheological properties
are non-Newtonian. Bijels stabilized by prolate and oblate particles also have different shear properties. For bijels stabilized by oblate particles, the stress curves
are more spread out suggesting that there is a larger variance in the properties arising from the change in the initial microstructure of the bijel. From the plot
the increase in particle order reduces the yield stress of the bijel. For bijels stabilized by rod like particles, the stress response of each processing 
condition is closer together, indicating a smaller variation in the response to the applied magnetic field compared to oblate particles. The shear response upon application
of the magnetic field also suggest that increased ordering of prolate particles makes the bijel more resistant to shear. We quantify the Herschel-Buckley fit parameters and 
tabulate the results in Table \ref{table:rheology_fit}.

\begin{table}[h!]
    \centering
    \renewcommand{\arraystretch}{1.5}  % Increases row height
    \begin{tabular}{||c c c c c||} 
     \hline
     Processing & $\alpha$ & $n$ & $K \frac{\Delta m (\Delta t)^{n-2}}{\Delta x} $ & $\sigma_{y} \frac{\Delta m}{\Delta x (\Delta t)^2}$ \\ [0.5ex] 
     \hline\hline
     $\bar{B}: 0 \rightarrow 0$ & 0.5 & 0.511 & 0.999 & $7.168 \cdot 10^{-5}$ \\ 
     \hline
     $\bar{B}: 0 \rightarrow 1$ & 0.5 & 0.579 & 0.944 & $4.008 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 0$ & 0.5 & 0.642 & 0.888 & $4.945 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 1$ & 0.5 & 0.637 & 0.914 & $1.715 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 0 \rightarrow 0$ & 2 & 0.561 & 0.965 & $4.618 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 0 \rightarrow 1$ & 2 & 0.609 & 0.944 & $4.101 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 0$ & 2 & 0.597 & 0.899 & $5.025 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 1$ & 2 & 0.606 & 0.885 & $6.435 \cdot 10^{-5}$ \\ [1ex] 
     \hline
    \end{tabular}
    \caption{Hershel Buckley fit parameters for different processing conditions applied to bijels stabilized by ellipsoidal particles.}
    \label{table:rheology_fit}
\end{table}

The results in Table \ref{table:rheology_fit} show that bijels remain shear thinning regardless of the processing history and particle morphology.
The particle morphology dependent differences in the shear response characterized in Figure \ref{fig:stress_strain} are quantified in Table \ref{table:rheology_fit}
where the yield stress differences predicted is clear to see. The bijel templates with an applied field of $\bar{B} = 1$ have larger flow indices
than bijel templates made with no applied field. The same trend exists here where the application of a magnetic field makes the bijel more Newtonian.
The yield stress of bijels stabilized with oblate particles decrease as the initial order of the particles and the applied field onto the bijel increase. 
The yield stress of bijels stabilized by prolate particles increases as the initial order of the particles and the applied field onto the bijel increase. In previous aims, we 
characterized that application of magnetic fields created orientational ordering of the bijels and that the local ordering of the bijels change. We  
showed that as magnetic fields were applied, oblate particles have lower local ordering on the interface while for prolate particles it increases.
Past investigations have identified that particle monolayers buckle more easily as local particle ordering decreases. \cite{prakash_buckling_2024} 
They identified that local ordering can enhance stress distribution among particles, making locally ordered particle monolayers better at withstanding shear.
We visualize the ordering of the particles at the interface to the z direction for both particle morphologies at $Ca_s = 10^{-6}$ for the 
case $\bar{B}: 1 \rightarrow 0$.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.4]{../figures/results/paper3/tz_concat_startB-1_endB-0.png} 
    \caption{Visualization of the particle monolayer of bijels stabilized by prolate and oblate ellipsoidal particles with processing history $\bar{B}: 1 \rightarrow 0$
             undergoing an applied shear of $Ca_s = 10^{-6}$. Particles with a deeper purple shade have a larger angle to the z axis. The top row shows the initial 
             configuration of the particles while the bottom row shows the final configuration of the particles. } 
    \label{fig:particle_viz_tz} 
\end{figure}

In Figure \ref{fig:particle_viz_tz}, the particles are initially ordered to the direction of the applied magnetic field. Upon application of shear and removal
of the magnetic field, We see that the particles begin reorienting away from the z direction. We also observe that the reorientation of particles uncovers the interface.
The reorientation observed is particle morphology specific with prolate particles seeing more reorientation away from the applied shear direction than oblate particles.
Looking at the images, we see that prolate particles rotate away from the direction of the magnetic field and applied shear towards the direction of the shear gradient,
while oblate particles reorient towards minimizing area to the applied shear direction. While these snapshots show the initial and final orientations of the particles at the
interface, the time evolution of the particle monolayer changes to identify its role in the rheological properties characterized.

\subsection{Particle properties}

Particles at the interface of bijels stabilized with ellipsoidal particles reorient at the interface to minimize interfacial and
steric energies. \cite{gunther_timescales_2014} The particles stabilizing bijels in confined systems under shear have been shown to
orient to the direction of shear and ejected from the interface at sufficiently long times or shear.\cite{bonaccorso_shear_2020} 
We probe these aspects of the particle monolayer in this section. First, we plot the orientation of the particle to each cartesian axis 
to identify if shear induced ordering is present.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/angle_to_cartesian-SS.png} 
    \caption{Comparisons of the initial and final particle orientations of the particles to each cartesian direction under
             four processing histories for bijels stabilized with oblate and prolate ellipsoids at multiple shear rates. We show
             shear dependent ordering of the particle monolayer if the initial particle monolayer begins with particle order.} 
    \label{fig:particle_orientation_cartesian_shear} 
\end{figure}

From Figure \ref{fig:particle_orientation_cartesian_shear} we see that with bijel templates simulated under $\bar{B} = 0$, the
particles adopt a random orientation. Upon application of shear, we see that there is little directional ordering to the 
shear direction in the simulation time observed. Upon application of the magnetic field in the $\bar{B}:0 \rightarrow 1$ case, 
the particles order to the direction of the magnetic field. Here we see that the prolate particles order more readily to the applied
field than the oblate particles, seen as a larger clustering of particles at the $\theta_z = 0$ corner of the triangle. In the 
$\bar{B}:1 \rightarrow 0$ case particle morphology dependent partial relaxation of the particles away from the magnetic field direction 
is observed. Directionality of the particles to the field is still observed, suggesting that the ordering of the particles is not
immediately removed once the field is switched off. In the processing history of $\bar{B}:1 \rightarrow 1$, particles are locked
into ordering to the magnetic field direction. However we also see that some particles are reoriented away from the field direction
due to the applied shear. Prolate particles have a smaller distribution of particle orientations than oblate particles, suggesting that
the initial orientation and ordering of particles to the shear applied underpin the response observed. As the particles are adsorbed on the 
interface, the orientation of the particles are influenced by a balance between magnetic forces from the field and capillary forces 
from the interface. We can qualitatively calculate the time evolution of the capillary interactions using the interfacial angle.

>>>>

The rheological properties of bijels are critical to their processability and structural fidelity across a variety of manufacturing techniques. In solvent transfer-induced phase separation 
(STrIPS), for example, the flow rate of the casting mixture and the co-solvent ratio play a key role in controlling the microstructure of the bijel and determining whether droplets or 
rope-like structures emerge during synthesis \cite{haase_continuous_2015, haase_situ_2016}. Similarly, in volatile-induced phase separation (VIPS), the viscosity of the solvent rich and poor 
phases determines the extent of phase separation before gelation or jamming occurs, ultimately governing final domain morphology \cite{wang_scalable_2020}. In homogenization-driven methods, 
the imposed shear rate directly influences the interfacial jamming dynamics and can either assist or disrupt bicontinuous structure formation depending on processing windows 
\cite{huang_bicontinuous_2017}. Understanding these rheological dependencies is essential for tuning bijel architectures for advanced applications.

While the microstructure and synthesis of bijels have been extensively investigated, their rheological behaviour remains comparatively underexplored. One intriguing feature is the formation of 
monogels, in which remixing the liquid phases of a bijel preserves the interfacial particle network due to persistent attractive interactions between particles \cite{sanz_colloidal_2009}. This 
phenomenon has been observed in bijels composed of lutidine/water and styrene/butadiene oligomer mixtures, but is notably in systems based on ethylene glycol and nitromethane 
\cite{sanz_colloidal_2009, bai_dynamics_2015, tavacoli_novel_2011}. Rheologically, bijels typically exhibit shear thinning and thixotropic responses under steady shear, indicative of dynamic 
restructuring and gradual recovery of the internal network \cite{macmillan_rheological_2019}. In confined geometries, computational studies have shown that shear induces elongation 
of domains followed by particle alignment along the flow direction, detachment from the interface, and eventual collapse of the bicontinuous structure \cite{bonaccorso_shear_2020}. More recently, 
bijels have been conceptualized as two-dimensional colloidal glasses embedded in a percolating three-dimensional network, exhibiting rheological behaviour that bridges colloidal gels and jammed 
particle layers \cite{ching_bijel_2022}. However, existing frameworks for bijel rheology focus on bijels stabilized by spherical, non-stimuli responsive particles, and do not yet capture the complex 
rheology arising from anisotropic particle shapes or magnetic interactions, which are increasingly relevant in functionalized bijel systems.

While the rheology of emulsions stabilized by spherical particles has been extensively characterized, ellipsoidal particles introduce additional complexity through interfacial locking and anisotropic drag, 
which can suppress coalescence and resist flow differently under shear \cite{madivala_self-assembly_2009, daware_emulsions_2015, sun_assembly_2013}. Under shear, ellipsoidal particles 
can promote shear banding, wherein particle-fluid interactions create localized variations in velocity profiles \cite{xu_relation_2013}. Furthermore, suspensions of magnetically responsive particles, 
such as ferrofluids, exhibit a pronounced increase in viscosity due to field-induced self-assembly that restricts fluid flow \cite{qiao_magnetorheological_2012}. 
Theoretical predictions and experimental observations have identified that orientational ordering of anisotropic particles in the direction of shear reduces viscosity by minimizing hydrodynamic drag and interfacial 
disruption, allowing the fluid to flow more easily along the direction of least resistance.\cite{vermant_flow-induced_2005, xu_relation_2013}.

In this study, we investigate the rheological response of bijels stabilized by magnetically responsive ellipsoidal particles subjected to constant shear. Using a hybrid Lattice Boltzmann-Molecular Dynamics 
(LBM-MD) simulation framework, we model a binary fluid system containing suspended magnetic ellipsoids, with shear applied via Lees-Edwards boundary conditions along one axis and periodic conditions along 
the remaining boundaries. Our results show that these bijels exhibit Herschel-Bulkley-type rheology, with yield stress and shear-thinning behavior modulated by particle orientation at the interface. 
Specifically, increasing orientational ordering of ellipsoidal particles under the applied magnetic field reduces hydrodynamic resistance, resulting in lower effective viscosity and a diminished shear-thinning 
response. Furthermore, we find that the yield stress is dependent on particle shape and magnetic alignment, as magnetic stimuli alter the number and orientation of frictional contacts within the interfacial 
monolayer. These findings highlight the tunability of bijel rheology via particle morphology and magnetic field application, offering potential design pathways for adaptive and reconfigurable soft materials.

@@

To probe the rheological behavior of bijels stabilized by ellipsoidal particles, we analyze their response under constant shear with particular focus on the influence of interfacial particle ordering 
and magnetic field application. The initial bijel structures are obtained from previous simulations, consisting of bicontinuous morphologies stabilized by either prolate or oblate magnetic 
ellipsoids at a particle volume fraction of $\phi_p = 0.15$, generated with and without an applied magnetic field of dimensionless strength $\bar{B} = 1$. To evaluate the combined and decoupled effects 
of field history and applied shear, we systematically vary the magnetic field configuration, yielding four distinct processing pathways: 
$\bar{B}:0 \rightarrow 0$, $\bar{B}:0 \rightarrow 1$, $\bar{B}:1 \rightarrow 0$, and $\bar{B}:1 \rightarrow 1$. These setups enable us to isolate the contributions of pre-aligned versus dynamically aligning 
particles during flow. Shear is imposed via a range of dimensionless shear capillary numbers, defined as $Ca_s = \frac{\dot{\gamma} R_s \eta_f}{\sigma}$, where $\dot{\gamma}$ is the applied shear rate, 
$R_s = 7.9$ is the equivalent spherical radius of the ellipsoids, $\eta_f$ is the dynamic viscosity of the continuous phase, and $\sigma$ is the interfacial tension. 
\cite{frijters_effects_2012, yang_capillary_2022} We explore five shear conditions spanning $Ca_s = 1 \times 10^{-4}$ to $1 \times 10^{-6}$, capturing the transition from strong deformation to 
near-equilibrium interfacial response.

@@

Particle ordering plays a critical role in determining the rheological behavior of suspensions, as aligned particles can more readily slide past one another 
under shear, delaying structural breakdown. In bijels, where particles are adsorbed at the fluid-fluid interface, this interfacial confinement may significantly 
influence the degree of slip and the onset of yielding. Prior studies have shown that bijels exhibit shear-thinning behavior consistent with Herschel-Bulkley rheology, 
indicating a non-Newtonian response governed by both interfacial structure and particle dynamics \cite{macmillan_rheological_2019}. To investigate these effects in 
the presence of magnetic fields and varying degrees of particle ordering, we apply constant shear to the bijel templates and plot the resulting time evolution of 
the shear stress, as shown in Figure~\ref{fig:stress_time}.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.5]{../figures/results/paper3/stress-time_compare.png} 
    \caption{Time evolution of the shear stress of bijels stabilized with ellipsoidal particles as a function of the applied strain rate as
             a function of the initial microstructure and applied magnetic field. The observed stress response is difficult to say when a steady
             state is reached.} 
    \label{fig:stress_time} 
\end{figure}

Figure \ref{fig:stress_time} demonstrates that a definitive steady-state shear stress is not reached within the simulated timeframes across all conditions. 
This absence of steady-state behavior is attributed to the continued coarsening of the bicontinuous domains and the progressive reorientation of interfacial 
particles under shear \cite{tavacoli_novel_2011, macmillan_rheological_2019}. Tavacoli et al.\ showed that mechanical perturbation of bijels, such as the 
insertion of a needle to measure yield stress, resulted in irreversible structural deformation, particularly in ethanediol/nitromethane systems 
\cite{tavacoli_novel_2011}. Similarly, Macmillan et al.\ observed that repeated shearing of bijels synthesized by direct mixing led to degraded mechanical 
performance over time, suggesting a link between structural evolution and rheological response \cite{macmillan_rheological_2019}. To characterize the rheology 
of these evolving systems, we identify pseudo-steady-state regions within the stress-time profiles, defined as intervals where the rate of change of shear 
stress is minimal and persists for the longest continuous duration. Within these regions, the stress response is fitted to the applied shear rate using a 
Herschel-Bulkley model:

\begin{equation}
    \sigma_{xz} = \sigma_{y} + K\dot{\gamma}^{n}    
\end{equation}

In the Herschel-Bulkley model, $\sigma_{xz}$ denotes the computed shear stress, $\sigma_y$ is the yield stress, $K$ represents the flow consistency index, 
and $n$ is the flow index characterizing the degree of shear thinning. In certain cases, simulations conducted at the lowest shear capillary number 
($Ca_s = 10^{-6}$) exhibited insufficient stress resolution or high variability, and were excluded from model fitting where appropriate. For each case, 
$\sigma_{xz}$ is extracted from the pseudo-steady-state region of the stres-time curves, while the corresponding shear rate $\dot{\gamma}$ is derived from 
the imposed capillary number, $Ca_s$. The resulting fits for the various bijel systems are presented in Figure~\ref{fig:stress_strain}.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.5]{../figures/results/paper3/stress_strain-all.png} 
    \caption{Fitting the experimental shear stress to the strain rate using the Herschel-Bulkley model. We characterize the bijels are 
             shear thinning in all cases and that the bijels become less shear thinning as the particle order increases. We also
             demonstrate that the particle ordering affects the yield stress differently for particles stabilized by oblate and 
             prolate particles.} 
    \label{fig:stress_strain} 
\end{figure}

Figure~\ref{fig:stress_strain} presents the shear stress $\sigma_{xz}$ as a function of shear rate $\dot{\gamma}$ for bijels stabilized by oblate ($\alpha = 0.5$) 
and prolate ($\alpha = 2.0$) ellipsoidal particles, fitted using the Herschel-Bulkley model. In both cases, the curves exhibit nonlinear scaling with $\dot{\gamma}$, 
confirming a non-Newtonian, shear-thinning response. For bijels stabilized by oblate particles, the stress-strain curves are more widely separated, indicating 
greater sensitivity to the initial microstructural configuration induced by magnetic field history. Specifically, increased orientational ordering—achieved through 
magnetic pre-alignment—correlates with a reduced yield stress, as observed by the downward shift of the stress curves. In contrast, bijels stabilized by prolate 
particles exhibit more closely clustered stress responses across different magnetic conditions, suggesting a reduced dependence on field history. Interestingly, 
for the prolate case, the stress response under persistent ordering ($\bar{B}:1 \rightarrow 1$) exceeds that of disordered systems, implying that aligned rod-like 
particles enhance interfacial resistance to shear. These trends highlight the contrasting rheological behaviors imparted by particle shape and the role of 
field-induced anisotropy in modulating mechanical response. Quantitative fit parameters obtained from the Herschel-Bulkley model are provided in 
Table~\ref{table:rheology_fit}.

\begin{table}[h!]
    \centering
    \renewcommand{\arraystretch}{1.5}  % Increases row height
    \begin{tabular}{||c c c c c||} 
     \hline
     Processing & $\alpha$ & $n$ & $K \frac{\Delta m (\Delta t)^{n-2}}{\Delta x} $ & $\sigma_{y} \frac{\Delta m}{\Delta x (\Delta t)^2}$ \\ [0.5ex] 
     \hline\hline
     $\bar{B}: 0 \rightarrow 0$ & 0.5 & 0.511 & 0.999 & $7.168 \cdot 10^{-5}$ \\ 
     \hline
     $\bar{B}: 0 \rightarrow 1$ & 0.5 & 0.579 & 0.944 & $4.008 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 0$ & 0.5 & 0.642 & 0.888 & $4.945 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 1$ & 0.5 & 0.637 & 0.914 & $1.715 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 0 \rightarrow 0$ & 2 & 0.561 & 0.965 & $4.618 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 0 \rightarrow 1$ & 2 & 0.609 & 0.944 & $4.101 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 0$ & 2 & 0.597 & 0.899 & $5.025 \cdot 10^{-5}$ \\
     \hline
     $\bar{B}: 1 \rightarrow 1$ & 2 & 0.606 & 0.885 & $6.435 \cdot 10^{-5}$ \\ [1ex] 
     \hline
    \end{tabular}
    \caption{Hershel Bulkley fit parameters for different processing conditions applied to bijels stabilized by ellipsoidal particles.}
    \label{table:rheology_fit}
\end{table}

The results presented in Table~\ref{table:rheology_fit} confirm that all bijel systems exhibit shear-thinning behavior, regardless of particle morphology or 
magnetic field history. The particle-shape-dependent differences in shear response observed in Figure~\ref{fig:stress_strain} are quantitatively reflected in 
the Hersche-Bulkley fit parameters, particularly in the yield stress $\sigma_y$ and flow index $n$. Notably, bijels synthesized with an applied magnetic field 
($\bar{B} = 1$) exhibit higher flow indices than those without, indicating a shift toward more Newtonian-like behavior under increased particle ordering. This 
trend is especially pronounced in bijels stabilized by oblate particles, where increased magnetic alignment leads to a progressive reduction in yield stress. 
In contrast, bijels stabilized by prolate particles show an opposite trend, with yield stress increasing alongside particle alignment and magnetic field 
application. These findings are consistent with earlier observations from Aim 1, which showed that magnetic fields promote local orientational ordering at the 
interface—enhancing ordering for prolate particles while reducing it for oblate particles. Prior studies have demonstrated that reduced local order in 
interfacial monolayers leads to structural buckling and early mechanical failure, whereas increased ordering enables more uniform stress distribution 
across particles, thereby enhancing resistance to deformation \cite{prakash_buckling_2024}.

\subsection{Shear induced particle rearrangement}

Given that magnetic fields can modify the local arrangement of ellipsoidal particles at fluid interfaces, understanding how this arrangement evolves under 
different processing conditions is essential for interpreting the mechanical 
behavior observed. In particular, the contrasting effects seen for prolate and oblate particles, where increased magnetic ordering leads to a larger
and smaller yield stress respectively. To quantitatively assess the local arrangement of particles at the interface, we compute the ensemble-averaged 
Steinhardt bond orientational order parameters $q_2$ and $q_6$. 
\cite{steinhardt_bond-orientational_1983, lechner_accurate_2008, mickel_shortcomings_2013}. These parameters characterize the local structural ordering by 
projecting the orientations of neighboring particles onto spherical harmonics of degree \(l\), and are defined as

\begin{equation}
q_l(i) = \left( \frac{4\pi}{2l+1} \sum_{m=-l}^{l} \left| \frac{1}{N(i)} \sum_{j=1}^{n(i)} Y_{lm}(\vec{r}_{ij}) \right|^2 \right)^{\frac{1}{2}} ,
\end{equation} 
%\TODO{update formula to the one actually used}

Here, \(n(i)\) denotes the number of nearest neighbors of particle \(i\), \(\vec{r}_{ij}\) is the vector connecting the centers of particles \(i\) and \(j\), 
and \(m\) is an integer ranging from \(-l\) to \(l\) that indexes the spherical harmonic components \(Y_{lm}\). Neighbor lists are typically defined using a 
fixed cutoff radius around each particle, identifying nearby particles as neighbors. We calculate the python package Freud to calculate the bond order parameters 
and use a Voronoi cell to determine the particle neighborhood rather than the standard distance criterion \cite{ramasubramani_freud_2020,mickel_shortcomings_2013}. 
The Voronoi-based method avoids the need for a predefined cutoff and yields more robust neighbor definitions, particularly in disordered systems. Bond 
orientational order parameters have been widely employed to study structural transitions in colloidal and ceramic systems, including crystallization, nucleation, 
and glass formation \cite{vagberg_glassiness_2011, besseling_three-dimensional_2007, schall_structural_2007, ozawa_jamming_2012}. $q_6$ is most commonly used 
to identify crystal structures by matching the calculated values to reference values for specific crystal structures such as fcc or hcp. We plot the change in 
$q_6$ of the microstructures as a function of the processing applied in Figure \ref{fig:q6_rheology_init}

\begin{figure} 
    \centering 
    \includegraphics[scale=0.4]{../figures/results/paper3/Q6-start.png} 
    \caption{Plot of the relationship between the local particle order $\langle Q6 \rangle$ at the first timestep after initiation of shear and 
             yield stress ($\sigma_y$) in bijels across magnetic field histories for oblate 
             ($\alpha = 0.5$) and prolate ($\alpha = 2.0$) particle-stabilized systems.} 
    \label{fig:q6_rheology_init} 
\end{figure}

Figure~\ref{fig:q6_rheology_init} shows that for oblate particles (left panel), $\langle Q_6 \rangle$ decreases progressively with increasing magnetic alignment, 
indicating a loss of local hexagonal
order at the interface. This trend is accompanied by a reduction in yield stress, suggesting that disrupted interfacial packing compromises the bijel's 
ability to resist deformation. These findings align with prior observations that reduced local order facilitates monolayer buckling and structural failure 
under shear. In contrast, bijels stabilized by prolate particles display the opposite behavior, characterized as an increase in $\langle Q_6 \rangle$ with
the particle order field history, signifying enhanced local ordering, and this increase correlates with a substantial rise in yield stress. The positive 
coupling between local particle order and mechanical strength in prolate systems supports the hypothesis that nematic alignment reinforces the interfacial 
network, distributing stress more effectively under shear. Together, these results demonstrate a morphology-dependent link between magnetic field-induced 
ordering and rheological response.

\subsection{Shear induced reorientation of particles}

Bonaccorso et al. identified that the spherical particle stabilizers in their bijel simulations oriented to the direction of shear before being
ejected from the interface. \cite{bonaccorso_shear_2020} Due to the shape of ellipsoidal particles, they adopt a preferred orientation which
maximizes the interface covered by the particle. \cite{davies_dipolar_2015, davies_interface_2014} Energy must be imparted to rotate the particles
out of the interface. Upon application of shear, the orientation of particles on the interface will be a function of the applied shear rate, 
applied magnetic field and the capillary forces of the interface on the particle. \cite{cao_modeling_2021, chhabra_drag_1999, brenner_stokes_1963, naga_capillary_2021}
Therefore, the degree of interface angle change observed can shed light onto the rheological 
differences of bijels stabilized by prolate and oblate particles as the shear stress determined is affected by the particle rotation on the interface.

We analyzed the angle between each particles symmetry axis and the local interface normal by generating a mesh of the interface using the marching 
cubes algorithm and, for each particle, identified the nearest mesh vertex (excluding those beyond the length of the particle's long axis). We then calculated 
the angle $\psi$ between the particle orientation vector \(\hat{\vec{o}}_i\) and the local surface normal. The time evolution of the average interface
angle of all particles $\langle \psi \rangle$, calculated as a numerical average of $\psi$ for all particles, is shown in Figure~\ref{fig:interface_angle_shear}.
